---
title: "PROJECT TITLE"
author: "NAME HERE"
date: "TODAY'S DATE"
output: github_document
---

## Section 1. Introduction


The dataset we are using is “TMBD 5000 Movie Dataset.” 

Originally, we found this set on kaggle.com. The data originates from “The Movie Database” API, it was collected from this original dataset.

The data’s variables: 
budget
genres
homepage
id
keywords
original_language
original_title
overview
popularity
production_companies
production_countries
release_date
revenue
runtime
spoken_languages
status
tagline
title
vote_average
Vote_count

Our research question would be, what variables correlate to a high revenue rate? 


## Section 2. Data analysis plan

The outcome would be revenue, and predictors would include but not be limited to the following variables: budget, production_companies, keywords, spoken_languages, original_language, release_date, and revenue, popularity.

```{r}
library(tidyverse)
```

```{r load-data, message=FALSE}
movies <- read_csv("tmdb_5000_movies.csv")
```

```{r}
movies %>%
  summarise(meanrev = mean(revenue)) 
```

This is the average revenue for all the movies in the dataset. We would consider a movie to have a high revenue if they are greater than the average. 

```{r}
movies %>%
  ggplot(mapping= aes(original_language, log(revenue))) +
  geom_boxplot() +
  coord_flip()
```


## Section 3. Data