---
title: "PROJECT TITLE"
author: "Zuck(R)berg"
date: "25 April 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Your project goes here! Before you submit, make sure your chunks are turned off with `echo = FALSE`. 

You can add sections as you see fit. Make sure you have a section called Introduction at the beginning and a section called Conclusion at the end. The rest is up to you!

### Introduction

### Stuff

```{r load-data, message=FALSE}
library(tidyverse)
library(dplyr)
library(broom)
library(stringr)
library(infer)

movies <- read_csv("/cloud/project/data/tmdb_5000_movies.csv")
movies <- movies %>%
  select(-id, -homepage, -original_title)
movies <- movies %>%
  filter(original_language == "en")
movies <- movies %>%
  filter(budget > 0) %>%
  filter(revenue > 0)

```

For our dataset of 5000 movies, we decided to remove the variables id, homepage, and original_title, since we felt that these variables did not contribute to our research question of what contributed to a "successful" movie. Additionally, since we found problems with the budget and revenue variables of non-American movies (not in US dollars), we decided to filter for only english movies as a proxy for American movies. (Since there is no definite way to filter for only American movies). 
We also filtered for movies that had a budget and revenue that were nonzero, this is so that our data would not be skewed by movies that were not intended to draw revenue or were not of the same budget caliber of the rest of the films. 

```{r popularity-explanations}
ggplot(data = movies, aes(y = popularity)) +
  geom_boxplot()

movies %>%
  select(title, popularity) %>%
  arrange(desc(popularity)) %>%
  head(10)
```

We are going to not look at popularity because we can't figure out where this came from,,, etc explain this.

```{r high-budget-list}
movies %>%
  select(title, budget) %>%
  arrange(desc(budget)) %>%
  head(10)
movies %>%
  distinct(production_companies)
```

### Creating a "Profitability" Variables

The first profitability variable is a ratio of revenue to budget, to show by what percentage a movie was "profitable."
This second profitability variable is a categorical variable that deciphers if the movie was profitable or not, based on the profitiability ratio. If the ratio was greater than 1, it was considered profitable. It will simply say "yes" for profitable, and "no" for non-profitable.

```{r}
movies <- movies %>% 
  mutate(pratio = revenue/budget) 
movies <- movies %>%
  mutate(profit = case_when(
    pratio > 1 ~ "yes",
    pratio <= 1 ~ "no"
  )
  )
movies <- movies %>%
  filter(pratio <= 200)

```

### Making a Categorical "spoken languages" Variable 

Here, we will be creating a new vategorical variable, that signifies if the only language spoken in the film is english, or not. If the only spoken language is english, the value will be "yes," if there are other languages spoken or english is not used, the value will be "no."

```{r spoken-languages}

movies <- movies %>%
  mutate(spokenlength = str_length(spoken_languages))

movies <- movies %>%
  mutate(english = case_when(
    spokenlength == 40 ~ "yes",
    spokenlength > 40 ~ "no",
    spokenlength < 40 ~ "no"
  ))

```


### Splitting date variable

```{r separate-date-variable}
movies <- movies %>%
  separate(release_date, c("year", "month", "day"), sep = "-")

movies <- movies %>%
  mutate(year = as.numeric(year), month = as.numeric(month), day = as.numeric(day))
```


### Genre trends

```{r genre-trends}
movies <- movies %>%
  mutate(horror = ifelse(str_detect(genres, "Horror"), "yes", "no"))

horror_only <- movies %>%
  filter(horror == "yes") %>%
  group_by(year) %>%
  mutate(m_pratio = median(pratio)) 

ggplot(horror_only, aes(year, m_pratio)) +
  geom_bar(stat="identity") +
  labs(title = "Change in Median Profitability Ratio of Horror Movies",
       x = "Year",
       y = "Median Profitability Ratio")
```

```{r horror}
horror_count <- horror_only %>%
  count(year)

ggplot(horror_count, aes(year, n)) +
  geom_bar(stat="identity") +
  labs(title = "Change in Number of Horror Movies",
       x = "Year",
       y = "# of Horror Movies")
```

```{r action}
movies <- movies %>%
  mutate(action = ifelse(str_detect(genres, "Action"), "yes", "no"))

action_only <- movies %>%
  filter(action == "yes") %>%
  group_by(year) %>%
  mutate(m_pratio = median(pratio)) 

ggplot(action_only, aes(year, m_pratio)) +
  geom_bar(stat="identity") +
  labs(title = "Change in Median Profitability Ratio of Action Movies",
       x = "Year",
       y = "Median Profitability Ratio")
```

```{r action-count}
action_count <- action_only %>%
  count(year)

ggplot(action_count, aes(year, n)) +
  geom_bar(stat="identity") +
  labs(title = "Change in Number of Action Movies",
       x = "Year",
       y = "# of Action Movies")
```

### Title and Profitability

```{r title-len}
movies <- movies %>%
  mutate(oneword = ifelse(!str_detect(title, " "), "yes", "no")) 

title_analysis <- movies %>%
  group_by(oneword) %>%
  summarise(median = median(pratio))


ggplot(title_analysis, aes(oneword, median)) +
  geom_bar(stat="identity") +
  labs(title = "Difference in Median Profitability Ratio Between One Word and Multi Word Titles",
       x = "One Word Title?",
       y = "Median Profitability Ratio")
```


### Making a variable for holiday releases

```{r holiday-variable}
movies <- movies %>%
  mutate(holiday_release = case_when(
    # Period between day before Christmas Eve and New Year's Eve (inclusive)
    month == 12 & day %in% c(23, 24, 25, 26, 27, 28, 29, 30, 31) ~ "yes",
    # New Year's Day (and the day after), flex period for Martin Luther King Jr. Day
    month == 1 & day %in% c(1, 2, 15, 16, 17, 18, 19, 20, 21) ~ "yes",
    # Valentine's Day (and the days before and after), flex period for Presidents' Day
    month == 2 & day %in% c(13, 14, 15, 16, 17, 18, 19, 20) ~ "yes",
    # Flex period for Memorial Day
    month == 5 & day %in% c(25, 26, 27, 28, 29, 30, 31) ~ "yes",
    # Independence Day (and the days before and after)
    month == 7 & day %in% c(3, 4, 5) ~ "yes",
    # Flex period for Labor Day
    month == 9 & day %in% c(1, 2, 3, 4, 5, 6, 7) ~ "yes",
    # Flex period for Thanksgiving, Veterans' Day (and the days before and after)
    month == 11 & day %in% c(10, 11, 12, 22, 23, 24, 25, 26, 27, 28) ~ "yes",
    TRUE ~ "no"
  ))
```

### Making a variable for likely sequels

```{r likely-sequel}
series <- movies$title %>%
  str_detect(":")

sequel2 <- movies$title %>%
  str_detect("2$")

sequel3 <- movies$title %>%
  str_detect("3")

movies <- movies %>%
  mutate(likely_sequel = case_when(
    series == TRUE ~ "yes",
    sequel2 == TRUE ~ "yes",
    sequel3 == TRUE ~ "yes",
    TRUE ~ "no"
  ))
```

### Making a variable for major production companies


```{r major-production-company}
universal_list <- movies$production_companies %>%
  str_detect("Universal Pictures")

paramount_list <- movies$production_companies %>%
  str_detect("Paramount Pictures")

warner_list <- movies$production_companies %>%
  str_detect("Warner Bros.")

disney_list <- movies$production_companies %>%
  str_detect("Walt Disney Pictures")

columbia_list <- movies$production_companies %>%
  str_detect("Columbia Pictures")

dreamworks_list <- movies$production_companies %>%
  str_detect("DreamWorks")

twentiethc_list <- movies$production_companies %>%
  str_detect("Twentieth Century Fox Film Corporation")

lionsgate_list <- movies$production_companies %>%
  str_detect("Lionsgate")

marvel_list <- movies$production_companies %>%
  str_detect("Marvel Studios")

pixar_list <- movies$production_companies %>%
  str_detect("Pixar Animation Studios")

movies <- movies %>%
  mutate(if_major = case_when(
    universal_list == TRUE ~ "yes",
    paramount_list == TRUE ~ "yes",
    warner_list == TRUE ~ "yes", 
    disney_list == TRUE ~ "yes",
    columbia_list == TRUE ~ "yes",
    dreamworks_list == TRUE ~ "yes",
    twentiethc_list == TRUE ~ "yes",
    lionsgate_list == TRUE ~ "yes",
    marvel_list == TRUE ~ "yes",
    pixar_list == TRUE ~ "yes",
    TRUE ~ "no"
  )) %>%
  mutate(major_productionco = case_when(
    universal_list == TRUE ~ "universal",
    paramount_list == TRUE ~ "paramount",
    warner_list == TRUE ~ "warner",
    disney_list == TRUE ~ "disney",
    columbia_list == TRUE ~ "columbia",
    dreamworks_list == TRUE ~ "dreamworks",
    twentiethc_list == TRUE ~ "20thc",
    lionsgate_list == TRUE ~ "lionsgate",
    marvel_list == TRUE ~ "marvel",
    pixar_list == TRUE ~ "pixar",
    TRUE ~ "other"
  ))
```

Major production companies according to Wikipedia are Universal Pictures (NBCUniversal), Paramount Pictures (Viacom), WarnerBros. Pictures (WarnerMedia), Walt Disney Pictures (Walt Disney Studios), Columbia Pictures (Sony Pictures) (in the format Major film studio unit (Studio parent) from https://en.wikipedia.org/wiki/Major_film_studio#Present_2).

### Making a variable for tagline length

```{r tagline-length}
movies <- movies %>%
  mutate(tag_length = str_length(tagline))
movies
```

### Making a preliminary linear model

We aim to find an ideal linear model to predict profitability ratio based on a variety of factors; before constructing our model, we will remove observations that lack values for these factors. 

```{r filter-data}
# filter for NAs (no NAs in holiday_release, likely_sequel, budget, if_major, major_productionco, vote_average, vote_count, english, horror )
movies_filt <- movies %>%
  filter(!is.na(tag_length))

movies_filt <- movies_filt %>%
  mutate(major_productionco = fct_relevel(major_productionco, "other"))
movies 
```

We filter the dataset to remove movies with NA values so that our backwards model selection will work (the only NAs in the values we are considering are for tagline length)

```{r preliminary-linear-model}
m_pratio <- lm(pratio ~ holiday_release + likely_sequel + budget + if_major + major_productionco + tag_length + english + runtime + action + horror + runtime + oneword + budget*if_major + likely_sequel*budget, data = movies_filt)
m_pratio

tidy(m_pratio) %>%
  select(term, estimate)

```


We fit a regression model to predict pulse rate based on holiday_release, likely_sequel, budget, if_major, major_productionco, tag_length, english, runtime, action, horror, oneword, the interaction between budget and likely_sequel, and the interaction between budget and if_major. We chose to examine the those interactions because whether a movie is a sequel (and thus has a known revenue for its precedents and some expectation that people will see it) may be related to its budget and also the relation between budget and having access to the resources we assume accompany a major production company may also be related in an interesting way.

Linear Model: 

Profitability Ratio = `r tidy(m_pratio)[1,2]` + `r tidy(m_pratio)[2,2]`(Holiday Release) + `r tidy(m_pratio)[3,2]`(Likely Sequel) + `r tidy(m_pratio)[4,2]`(Budget) + `r tidy(m_pratio)[5,2]`(Major Production Company) + `r tidy(m_pratio)[6,2]`(20th Century Fox) `r tidy(m_pratio)[7,2]`(Columbia) + + `r tidy(m_pratio)[8,2]`(Disney) + `r tidy(m_pratio)[9,2]`(DreamWorks) + `r tidy(m_pratio)[10,2]`(Lionsgate) + `r tidy(m_pratio)[11,2]`(Marvel) + `r tidy(m_pratio)[12,2]`(Paramount) + `r tidy(m_pratio)[13,2]`(Pixar) + `r tidy(m_pratio)[14,2]`(Universal) + `r tidy(m_pratio)[15,2]`(Tagline Length) + `r tidy(m_pratio)[16,2]`(Spoken Language English) + `r tidy(m_pratio)[17,2]`(Runtime) + `r tidy(m_pratio)[18,2]`(Action) + `r tidy(m_pratio)[19,2]`(Horror) + `r tidy(m_pratio)[20,2]`(One Word Title) + `r tidy(m_pratio)[21,2]`(Budget * Major Production Company) + `r tidy(m_pratio)[22,2]`(Likely Sequel * Budget)

```{r r-orig-model}
glance(m_pratio)

rsq <- glance(m_pratio)$adj.r.squared
```

We compute the adjusted R-squared value for the filtered model for pulse as `r rsq`; this is a helpful measure of the goodness of the fit of the model.

```{r aic-model-selection}
selected_m_pratio <- step(m_pratio, direction = "backward")

tidy(selected_m_pratio) %>% 
  select(term, estimate)
```

We perform backwards model selection based on AIC, Akaike Information Criterion (another measure of the fit of a model), and obtain the best fit.

Selected Linear Model: 

Profitability Ratio = `r tidy(selected_m_pratio)[1,2]` + `r tidy(selected_m_pratio)[2,2]`(Holiday Release) + `r tidy(selected_m_pratio)[3,2]`(Likely Sequel) + `r tidy(selected_m_pratio)[4,2]`(Budget) + `r tidy(selected_m_pratio)[5,2]`(Major Production Company) + `r tidy(selected_m_pratio)[6,2]`(Spoken Language English) + `r tidy(selected_m_pratio)[7,2]`(Runtime) + `r tidy(selected_m_pratio)[8,2]`(Action) + `r tidy(selected_m_pratio)[9,2]`(Horror) + `r tidy(m_pratio)[10,2]`(One Word Title) + `r tidy(m_pratio)[11,2]`(Budget * Major Production Company) + `r tidy(m_pratio)[12,2]`(Likely Sequel * Budget)

```{r aic-model-comparison}
glance(m_pratio)$AIC
glance(selected_m_pratio)$AIC
```

Our AIC value decreases for the selected model, which indicates that it is a better fit. Thus factors of holiday_release, likely_sequel, budget, ETC TYPE THE REST HERE are the best predictors of profitability. 

```{r rsq-selected-model}
newrsq <- glance(selected_m_pratio)$r.squared
newrsq
```

Our R-squared value of `r newrsq` for our final selected model indicates that `r newrsq *100`% of the variation in Profitability can be well-explained by a linear relationship with INSERT LIST HERE


### Conclusion


