---
title: "PROJECT TITLE"
author: "Zuck(R)berg"
date: "25 April 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Your project goes here! Before you submit, make sure your chunks are turned off with `echo = FALSE`. 

You can add sections as you see fit. Make sure you have a section called Introduction at the beginning and a section called Conclusion at the end. The rest is up to you!

### Introduction

### Stuff

```{r load-data, message=FALSE}
library(tidyverse)

movies <- read_csv("/cloud/project/data/tmdb_5000_movies.csv")
movies <- movies %>%
  select(-id, -homepage, -original_title)
movies 
movies <- movies %>%
  filter(original_language == "en")

movies
```

```{r popularity-explanations}
ggplot(data = movies, aes(y = popularity)) +
  geom_boxplot()

movies %>%
  select(title, popularity) %>%
  arrange(desc(popularity)) %>%
  head(10)
```

We are going to not look at popularity because we can't figure out where this came from,,, etc explain this.

```{r high-budget-list}
movies %>%
  select(title, budget) %>%
  arrange(desc(budget)) %>%
  head(10)
movies %>%
  distinct(production_companies)
```


### Conclusion


